---
title: "Homework 2"
author: Yuqi Wang
output: github_document
---

This is my question for HW2.

```{r}
library(tidyverse)
library(readxl)
```

## Problem 1

First, define a path to the dataset.
```{r}
path_to_data = './data/Trash-Wheel-Collection-Totals-8-6-19.xlsx'
```

Read the Mr. Trashwheel dataset.
```{r}
trashwheel = 
  read_excel(path = path_to_data,
             sheet = "Mr. Trash Wheel",
             range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls),
         sports_balls = as.integer(sports_balls))
# the default sheet for read_excel() is the first sheet in the excel.
# if we use round(), the type of the variable will still be double, but if we use as.integer, the type will be integer after transformation.
```

Read 2018 and 2017 precipitation data.

```{r}
precip_2018 = 
  read_excel('./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx',
             sheet = "2018 Precipitation",
             skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel('./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx',
             sheet = "2017 Precipitation",
             skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now, combine 2017-2018 precipitation data.

```{r}
month_df = tibble(
  month = 1:12,
  month_name = month.name
)

precip_data = 
  bind_rows(precip_2018, precip_2017) #%>% 
  #mutate(month = month.name[month])

left_join(precip_data, month_df, by = 'month')
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel)` rows in our final dataset. Additional data sheets include month precipitation data in 2017 and 2018.