---
title: "Homework 2"
author: Yuqi Wang
output: github_document
---

This is my question for HW2.

```{r}
library(tidyverse)
library(readxl)
```

## Problem 1

First, define a path to the dataset.
```{r}
path_to_data = './data/Trash-Wheel-Collection-Totals-8-6-19.xlsx'
```

Read the Mr. Trashwheel dataset.
```{r}
trashwheel = 
  read_excel(path = path_to_data,
             sheet = "Mr. Trash Wheel",
             range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls),
         sports_balls = as.integer(sports_balls))
# the default sheet for read_excel() is the first sheet in the excel.
# if we use round(), the type of the variable will still be double, but if we use as.integer, the type will be integer after transformation.
```

Read 2018 and 2017 precipitation data.

```{r}
precip_2018 = 
  read_excel('./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx',
             sheet = "2018 Precipitation",
             skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel('./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx',
             sheet = "2017 Precipitation",
             skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now, combine 2017-2018 precipitation data.

```{r}
month_df = tibble(
  month = 1:12,
  month_name = month.name
)

precip_data = 
  bind_rows(precip_2018, precip_2017) #%>% 
  #mutate(month = month.name[month])

left_join(precip_data, month_df, by = 'month')
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel)` rows in our final dataset. Additional data sheets include month precipitation data in 2017 and 2018.

## Problem 2

First, read the data set, clean the names, and select the required columns.
Next, convert the entry variable to a logical variable.

```{r}
nyc_transit_df = read.csv('./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv', na.strings = c("","NA")) %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, 'YES' = TRUE, 'NO' = FALSE)) %>% 
  distinct()
nyc_transit_df
```

########should I remove all the repeated entry?


Description of the dataset:
This data set includes information about the NYC subway transits. It contains the following variables: `r ls(nyc_transit_df)`. So far, the data has been imported into R, and the variable names have been cleaned, and the entry variable was recoded as a logical variable. The size of the data set is `r nrow(nyc_transit_df)` rows x `r ncol(nyc_transit_df)` columns.

The data is not tidy because from route1 to route11 we see that the route information is spread across 11 columns.

#######the columns represent the variables, the rows represent different observations, and each value has a cell.

Next, answering the three questions:

```{r}
# Question 1: distinct stations
station_num = distinct(nyc_transit_df, line, station_name)

# Question 2: ADA compliant stations
ada_num = distinct(nyc_transit_df, line, station_name, ada) %>% 
  filter(ada == TRUE)

# Question 3: proportion of entrances without vending
no_vending_num = nyc_transit_df %>% 
  select(line, station_name, vending) %>% 
  filter(vending == 'NO')

no_vending_prop = nrow(no_vending_num)/nrow(nyc_transit_df)


#prop.table(table(data$vending))
#data %>% 
  #group_by(vending) %>% 
  #summarise(prop = n()/nrow(data))

#nrow(distinct(filter(nyc_transit_df, vending == 'NO'), line, station_name, keep_all = FALSE))/nrow(distinct(nyc_transit_df, line, station_name, keep_all = FALSE))
```


* There are `r nrow(station_num)` distinct stations.
* There are `r nrow(distinct(filter(nyc_transit_df, ada == TRUE), line, station_name, keep_all = FALSE))` ADA compliant stations.
* There are `r nrow(ada_num)` ADA compliant stations.


* The proportion of station entrances/exits without vending allow entrance is `r no_vending_prop`

`r nrow(distinct(filter(nyc_transit_df, vending == 'NO'), line, station_name))/nrow(distinct(nyc_transit_df, line, station_name))`.

Next, separate route name and route number as separate variables.

```{r}
nyc_transit = nyc_transit_df %>% 
  mutate(route8 = as.character(route8), 
        route9 = as.character(route9),
        route10 = as.character(route10),
        route11 = as.character(route11))

nyc_transit_tidy = nyc_transit %>% 
  pivot_longer(
    route1:route11,
    names_to = "route",
    names_prefix = "route",
    values_to = "route_name"
  ) %>% 
  drop_na(route_name)

# remove NA data or " " data for route

```

Select distinct station that serves A train

```{r}
a_train_serv = nyc_transit_tidy %>% 
  filter(route_name == "A") %>% 
  distinct(line, station_name, ada)

a_train_ada = a_train_serv %>% 
  filter(ada == TRUE)
```

There are `r nrow(a_train_serv)` distinct stations serving A train.
Among these stations, `r nrow(a_train_ada)` are ADA compliant.


## Problem 3

First step: process the pols_mon data
```{r}
#month_df = tibble(
  #month = 01:12,
  #month_name = month.name
#)
pols_month = read.csv('./data/fivethirtyeight_datasets/pols-month.csv') %>% 
  separate(mon, c("year", "month", "day"), "-") %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day)) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(president = prez_gop) %>% 
  select(-day, -prez_gop, -prez_dem) %>% 
  relocate(year, month, president)

# president = 0 means dem, =1 means gop
  #left_join(., month_df, by = 'month') %>% 
```

Second step: processing the snp data
```{r}
snp = read.csv('./data/fivethirtyeight_datasets/snp.csv') %>% 
  separate(date, c("month", "day", "year"), "/") %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day)) %>% 
  mutate(month = month.name[month]) %>% 
  select(-day) %>% 
  relocate(year, month)
```

Third step: tidying the unemployment data
```{r}
snp = read.csv('./data/fivethirtyeight_datasets/snp.csv') %>% 
  separate(date, c("month", "day", "year"), "/") %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day)) %>% 
  mutate(month = month.name[month]) %>% 
  select(-day) %>% 
  relocate(year, month)
```


